---
title: "C1F-LR"
output: pdf_document
date: "2024-05-19"
---

# 0. Set-up
```{r set up, include=FALSE}
#Library
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggmosaic))
suppressPackageStartupMessages(library(modelr))
suppressPackageStartupMessages(library(boot))
suppressPackageStartupMessages(library(infer))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(caret))
```

# 1. Train/Test dataset
```{r}
# File
file_path <- "AgeDataset.csv"

# Data set
df <- read.csv(file_path)

# Filter (only 'Male' or 'Female')
df_filtered <- df %>% filter(Gender %in% c("Male", "Female"))

# Set seed for reproducibility
set.seed(42)

# Split (3:7)
trainIndex <- createDataPartition(df_filtered$Gender, p = 0.7, list = FALSE, times = 1)
train_df <- df_filtered[trainIndex, ]
test_df <- df_filtered[-trainIndex, ]

# Save to CSV files
write.csv(train_df, "train_age.csv", row.names = FALSE)
write.csv(test_df, "test_age.csv", row.names = FALSE)

# Check
print("Training and testing datasets have been saved successfully.")
```

# 2. Tidying the dataset
```{r}
#takeout-train
train_df <- train_df %>% select(- Short.description) 

#takeout-test
test_df <- test_df %>% select(- Short.description) 
```
```{r}
#train grouping 
train_age <- train_df %>%
  mutate(Contains.four = ifelse(grepl("4", as.character(Birth.year)), 1, 0)) %>% 
  mutate(Contains.super = as.logical(Contains.four)) %>%
  filter(Birth.year > 0)

#test grouping
test_age <- test_df %>%
  mutate(Contains.four = ifelse(grepl("4", as.character(Birth.year)), 1, 0)) %>%
  mutate(Contains.super = as.logical(Contains.four)) %>%
  filter(Birth.year > 0)
```

# 3. Trained Graph

## 3.1. Quantitative Vari Graph

### 3.1.1. percentaged
```{r}
train_age %>% 
  pivot_longer(cols = Age.of.death|Birth.year,
               names_to = "info1",
               values_to = "value1") %>% 
  ggplot() +
  geom_histogram(
    mapping = aes(
      x = value1,
      y = after_stat(count) / sum(after_stat(count)) * 100,
      fill = Contains.super
    ),
    position = "identity",
    alpha = 0.6
  ) +
  facet_wrap(~info1, scales = "free") +
  labs(
    x = "value",
    y = "frequency",
    title = "Histogram of superstition due to age of death and birth year"
  )
```

### 3.1.2. not-percentaged
```{r}
train_age %>% 
  pivot_longer(cols = Age.of.death|Birth.year,
               names_to = "info1",
               values_to = "value1") %>% 
  ggplot() +
  geom_histogram(
    mapping = aes(
      x = value1,
      fill = Contains.super
    ),
    position = "identity",
    alpha = 0.6
  ) +
  facet_wrap(~info1, scales = "free") +
  labs(
    x = "value",
    y = "frequency",
    title = "Histogram of superstition due to age of death and birth year"
  )
```

## 3.2. Qualitative Vari Graph

### 3.2.1. Top-filtered
```{r}
# Count and get the top 10 Occupations
top_occupations <- train_age %>%
  count(Occupation) %>%
  top_n(10, wt = n) %>%
  pull(Occupation)

# Count and get the top 10 Countries
top_countries <- train_age %>%
  count(Country) %>%
  top_n(10, wt = n) %>%
  pull(Country)

# Filter the train_age dataframe based on top Occupations and Countries
top_filtered <- train_age %>%
  filter(Occupation %in% top_occupations) %>% 
  filter(Country %in% top_countries)
```

### 3.2.2. percentaged
```{r}
# Pivot the dataset and plot
top_filtered %>%
  pivot_longer(cols = c(Country, Occupation, Gender), names_to = "info2", values_to = "value2") %>%
  ggplot() +
  geom_bar(
    mapping = aes(
      x = value2, 
      y = after_stat(count) / sum(after_stat(count)) * 100,
      fill = Contains.super), alpha = 0.6) +
  facet_wrap(~info2, scales = "free") +
  labs(
    x = "value", 
    y = "frequency", 
    title = "Bar graph of superstition due to occupation,gender, and country"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 3.2.3. not-percentaged
```{r}
# Pivot the dataset and plot
top_filtered %>%
  pivot_longer(cols = c(Country, Occupation, Gender), names_to = "info2", values_to = "value2") %>%
  ggplot() +
  geom_bar(
    mapping = aes(
      x = value2, 
      fill = Contains.super), alpha = 0.6) +
  facet_wrap(~info2, scales = "free") +
  labs(
    x = "value", 
    y = "frequency", 
    title = "Bar graph of superstition due to occupation,gender, and country"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 3.3 Mosaic Graph
```{r}
# Filter out rows with missing values in the relevant columns
train_imputed <- top_filtered %>% 
  filter(!is.na(Gender) & !is.na(Occupation) & !is.na(Contains.super))

# Convert necessary columns to factors
train_imputed <- train_imputed %>%
  mutate(Gender = as.factor(Gender),
         Occupation = as.factor(Occupation),
         Contains.super = as.factor(Contains.super))
```
```{r}
# Create the mosaic plot
train_imputed %>% 
  ggplot() +
  geom_mosaic(aes(x = product(Gender, Occupation), fill = Gender)) +
  facet_grid(. ~ Contains.super, scales = "free") +
  labs(x = "Occupation", 
       y = "Gender", 
       title = "Mosaic Plot of Superstition") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

# 4. Logistic Regression

### 4.1. Women LR 

## 4.1.1. Women-Manipulate the data
```{r}
# Count and get the top 10 Occupations
top_occupations_2 <- test_age %>%
  count(Occupation) %>%
  top_n(10, wt = n) %>%
  pull(Occupation)

# Count and get the top 10 Countries
top_countries_2 <- test_age %>%
  count(Country) %>%
  top_n(10, wt = n) %>%
  pull(Country)

# Filter the train_age dataframe based on top Occupations and Countries
top_filtered_2 <- test_age %>%
  filter(Occupation %in% top_occupations_2) %>% 
  filter(Country %in% top_countries_2)
```
```{r}
# Filter out rows with missing values in the relevant columns
test_imputed <- top_filtered_2 %>% 
  filter(!is.na(Gender) & !is.na(Occupation) & !is.na(Contains.super))

# Convert necessary columns to factors
test_imputed <- test_imputed %>%
  mutate(Gender = as.factor(Gender),
         Occupation = as.factor(Occupation),
         Contains.super = as.factor(Contains.super))
```
```{r}
women_train_age <- train_imputed %>% 
    filter(Gender == 'Female') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))

women_test_age <- test_imputed %>% 
    filter(Gender == 'Female') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))
```

### 4.1.2. modeling/pred - train
```{r}
w_train_model <- glm(Contains.super ~ Age.of.death, data = women_train_age, family = "binomial")
```
```{r}
w_train_preds <-  women_train_age %>% 
  add_predictions(w_train_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

w_train_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 4.1.3 cross validation1 - train
```{r}
# Cost function for cross validation
cost <- function(obs, pred) {
  outcome <- pred > 0.5
  return(1 - (sum(obs == outcome) / length(obs)))
}

logistic_sv1 <- cv.glm(women_train_age, w_train_model, cost, K = 5)

logistic_sv1$delta
```

### 4.1.4. modeling/pred - test
```{r}
w_test_model <- glm(Contains.super ~ Age.of.death + Occupation + Country, 
                    data = women_test_age, family = "binomial")
```
```{r}
w_test_preds <-  women_test_age %>% 
  add_predictions(w_test_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

w_test_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 4.1.5. cross validation2 - test
```{r}
# Cost function for cross validation
cost <- function(obs, pred) {
  outcome <- pred > 0.5
  return(1 - (sum(obs == outcome) / length(obs)))
}

logistic_sv2 <- cv.glm(women_test_age, w_test_model, cost, K = 5)

logistic_sv2$delta
```

### 4.2. Men LR 

## 4.2.1.  Men-Manipulate the data
```{r}
men_train_age <- train_imputed %>% 
    filter(Gender == 'Male') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))

men_test_age <- test_imputed %>% 
    filter(Gender == 'Male') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))
```


### 4.2.2. modeling/pred - train
```{r}
m_train_model <- glm(Contains.super ~ Age.of.death, data = men_train_age, family = "binomial")
```
```{r}
m_train_preds <-  men_train_age %>% 
  add_predictions(m_train_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

m_train_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 4.2.3. cross validation3 - train
```{r}
logistic_sv3 <- cv.glm(men_train_age, m_train_model, cost, K = 5)

logistic_sv3$delta
```

### 4.2.4. modeling/pred - test
```{r}
m_test_model <- glm(Contains.super ~ Age.of.death + Occupation + Country,
                    data = men_test_age, family = "binomial")
```
```{r}
m_test_preds <-  men_test_age %>% 
  add_predictions(m_test_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

m_test_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 4.2.5. cross validation3 - test
```{r}
logistic_sv4 <- cv.glm(men_test_age, m_test_model, cost, K = 5)

logistic_sv4$delta
```

