---
title: "CDS 101 Final Project"
author: "Group 4"
date: "`r Sys.Date()`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    highlight: tango
    toc: false
    df_print: kable
    fig_caption: no
    number_sections: no
    dev: pdf
---

# 0. Set-up & Introduction
```{r set up, include=FALSE}
#Library
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggmosaic))
suppressPackageStartupMessages(library(modelr))
suppressPackageStartupMessages(library(boot))
suppressPackageStartupMessages(library(broom))
options(tigris_use_cache = TRUE)
suppressPackageStartupMessages(library(infer))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(caret))

#Data set
df <- read_csv("AgeDataset.csv")

#Helper functions
#load("bootstrap_cohens_d.RData")
```
>You can write introduction here.

# 1. Manipulating dataset
```{r}
#Rename
df_renamed <- df %>%
  rename(
    Birth.year = 'Birth year',
    Death.year = 'Death year',
    Manner.of.death = 'Manner of death',
    Age.of.death = 'Age of death'
  )

#Missing value
df_renamed[
  is.na(df_renamed$Birth.year) | df_renamed$Birth.year == "" |
  is.na(df_renamed$Death.year) | df_renamed$Death.year == "" |
  is.na(df_renamed$Country) | df_renamed$Country == "" |
  is.na(df_renamed$Occupation) | df_renamed$Occupation == "" |
  is.na(df_renamed$Manner.of.death) | df_renamed$Manner.of.death == "",
  c("Birth.year", "Death.year", "Country", "Occupation", "Manner.of.death")
] <- NA
```
```{r}
#Unnecessary
df_renamed <- df_renamed %>% 
  select(- `Short description`) 

#Filter (only 'Male' or 'Female')
df_filtered <- df_renamed %>% filter(Gender %in% c("Male", "Female"))

#Mutate group 4/0
AgeDataset <- df_filtered %>%
  mutate(Contains.four = ifelse(grepl("4", as.character(Birth.year)), 1, 0)) %>% 
  mutate(Contains.super = as.logical(Contains.four)) %>%
  filter(Birth.year > 0)
```
> We renamed the variables and filtered the gender into only two due to the complexity. More, we add two variable that shows number or status (true/false) of either they have number 4 in their birth year or not.

# 2. Train/Test dataset
```{r}
# Set seed for reproducibility
set.seed(42)

# Split (3:7)
trainIndex <- createDataPartition(AgeDataset$Gender, p = 0.7, list = FALSE, times = 1)
train_age <- AgeDataset[trainIndex, ]
test_age <- AgeDataset[-trainIndex, ]

# Save to CSV files
write.csv(train_age, "train_age.csv", row.names = FALSE)
write.csv(test_age, "test_age.csv", row.names = FALSE)

# Check
print("Training and testing datasets have been saved successfully.")
```
> For the later part, logistic regression, we divded into 2 dataset, train_age & test_age.

# 3. EDA
```{r}
AgeDataset_grouped <- AgeDataset %>%
  group_by(Contains.four, Gender, Occupation) %>%
  summarize(count = n(), .groups = 'drop')
```

## 3.1. Occupation

### 3.1.1 Filtering the top 20 Occupation
```{r}
AgeDataset_oc_count <- AgeDataset %>%
	count(Occupation)
```
```{r}
selected_occupations <- c("Artist", "Politician", "Athlete", "Researcher", "Military personnel",
                          "Religious figure", "Businessperson", "Architect", "Journalist", "Teacher",
                          "Physician", "Engineer", "Judge", "Lawyer", "Jurist", "Aristocrat", 
                          "Entrepreneur", "Philosopher", "Translator")
```

### 3.1.2. Occupation Visualization
```{r}
AgeDataset_occupation <- AgeDataset %>%
mutate(most = case_when(Occupation %in% selected_occupations ~ 1, TRUE ~ 0)) %>%
 	select(Id, Birth.year, Occupation, Contains.four, most)
```

```{r}
oc_filtered_data <- AgeDataset_occupation[AgeDataset_occupation$most == 1, ]
```

```{r}
oc_filtered_data %>%
  ggplot() +
  geom_bar(mapping = aes(x = Occupation, 
                         fill = factor(Contains.four)), 
           position = "stack") +
  labs(title = "Top 20 Occupation", x = "Occupation", y = "Count") +
 	theme_minimal() +
 	scale_fill_discrete(name = "Contains Four")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## 3.2. Manner of death
```{r}
AgeDataset_mod <- AgeDataset %>%
  select(Id, Birth.year, Manner.of.death, Contains.four)
```

### 3.2.1 Cleaning the Data
```{r}
mod_mutated <- AgeDataset_mod %>%
  mutate(clean_mod = 
ifelse(grepl("capital punishment", Manner.of.death), "capital punishment",
ifelse(grepl("homicide", Manner.of.death), "homicide",
ifelse(grepl("accident", Manner.of.death), "accident",
ifelse(grepl("natural causes", Manner.of.death), "natural causes",
ifelse(grepl("suicide", Manner.of.death), "suicide",
ifelse(grepl("unnatural death", Manner.of.death), "unnatural death",
ifelse(grepl("war", Manner.of.death), "war", Manner.of.death))))))))

```

### 3.2.2. Filtering the top 20 MOD
```{r}
mod_counted <- mod_mutated %>%
  count(clean_mod)
```

```{r}
selected_mod <- 
  c("natural causes", "suicide", "accident", "homicide", "capital punishment", 
    "unnatural death", "death in battle", "war", "extrajudicial killing")
```

### 3.2.3. MOD Visualization
```{r}
mod_selected <- mod_mutated %>%
            mutate(top10 = case_when(Manner.of.death %in% selected_mod ~ 1, TRUE ~ 0))

```

```{r}
filtered_mod <- mod_selected[mod_selected$top10 == 1, ]
```

```{r}
filtered_mod %>%
  ggplot() +
  geom_bar(mapping = aes(x = clean_mod,
                       fill = factor(Contains.four)),
           position = "stack") +
  labs(title = "Top 10 Manner of Death", x = "Manner of Death", y = "Count") +	
  theme_minimal() +
  scale_fill_discrete(name = "Contains Four")  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# 4. Hypothesis test

> Since female and male does not have different value of the average life span, out team decided to do the hypothesis test 2 times for each gender. The other genders are not considered due to the complexity. This hypothesis is not based on the either train_age or test_age but the total dataset.

# 4.1. Women

## 4.1.1. Filter the data
```{r}
female_data <- AgeDataset %>%
  filter(Gender == 'Female')

female_data <- female_data %>%
  mutate(Contains.super = as.factor(Contains.super))

levels(female_data$Contains.super)
```

## 4.1.2. Null distribution
```{r}
set.seed(123)
f_null <- female_data %>%
  specify(formula = Age.of.death ~ Contains.super) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
```
> Why not "true"/"false" but "1", "0"?

## 4.1.3. Observed difference
```{r}
f_observed_diff <- female_data %>%
  specify(response = Age.of.death, explanatory = Contains.super) %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
```

## 4.1.4. P-value & visualization
```{r}
f_p_value <- f_null %>%
  get_p_value(obs_stat = f_observed_diff, direction = "both")

f_null %>%
  visualize() +
  shade_p_value(obs_stat = f_observed_diff, direction = "both") +
  labs(title = "Simulation Based Null Distribution",
       x = "Difference in Means",
       y = "Count")
```
```{r}
print(f_p_value)
```

> Can it be different result?
> We should reject the null distribution, therefore the lifespan is different.

# 4.2. Men

## 4.2.1. Filter the data
```{r}
male_Data <- AgeDataset %>%
  filter(Gender == 'Male')

male_Data <- male_Data %>%
  mutate(Contains.super = as.factor(Contains.super))

levels(male_Data$Contains.super)
```

## 4.2.2. Null distribution
```{r}
set.seed(999)
m_null <- male_Data %>%
  specify(formula = Age.of.death ~ Contains.super) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
```

## 4.2.3. Observed difference
```{r}
m_observed_diff <- male_Data %>%
  specify(Age.of.death ~ Contains.super) %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
```

## 4.2.4. P-value & visualization
```{r}
m_p_value <- m_null %>%
  get_p_value(obs_stat = m_observed_diff, direction = "both")

m_null %>%
  visualize() +
  shade_p_value(obs_stat = m_observed_diff, direction = "both") +
  labs(title = "Visualization Null Distribution for Male Lifespan",
       x = "Difference in Means",
       y = "Count")
```
```{r}
print(m_p_value)
```
> For male... it is not right?

# 5. Trained Graph

## 5.1. Quantitative Vari Graph

### 5.1.1. percentaged
```{r}
train_age %>% 
  pivot_longer(cols = Age.of.death|Birth.year,
               names_to = "info1",
               values_to = "value1") %>% 
  ggplot() +
  geom_histogram(
    mapping = aes(
      x = value1,
      y = after_stat(count) / sum(after_stat(count)) * 100,
      fill = Contains.super
    ),
    position = "identity",
    alpha = 0.6
  ) +
  facet_wrap(~info1, scales = "free") +
  labs(
    x = "value",
    y = "frequency",
    title = "Histogram of superstition due to age of death and birth year"
  )
```

### 5.1.2. not-percentaged
```{r}
train_age %>% 
  pivot_longer(cols = Age.of.death|Birth.year,
               names_to = "info1",
               values_to = "value1") %>% 
  ggplot() +
  geom_histogram(
    mapping = aes(
      x = value1,
      fill = Contains.super
    ),
    position = "identity",
    alpha = 0.6
  ) +
  facet_wrap(~info1, scales = "free") +
  labs(
    x = "value",
    y = "frequency",
    title = "Histogram of superstition due to age of death and birth year"
  )
```

## 5.2. Qualitative Vari Graph

### 5.2.1. Top-filtered
```{r}
# Count and get the top 10 Occupations
top_occupations <- train_age %>%
  count(Occupation) %>%
  top_n(10, wt = n) %>%
  pull(Occupation)

# Count and get the top 10 Countries
top_countries <- train_age %>%
  count(Country) %>%
  top_n(10, wt = n) %>%
  pull(Country)

# Filter the train_age dataframe based on top Occupations and Countries
top_filtered <- train_age %>%
  filter(Occupation %in% top_occupations) %>% 
  filter(Country %in% top_countries)
```

### 5.2.2. percentaged
```{r}
# Pivot the dataset and plot
top_filtered %>%
  pivot_longer(cols = c(Country, Occupation, Gender), names_to = "info2", values_to = "value2") %>%
  ggplot() +
  geom_bar(
    mapping = aes(
      x = value2, 
      y = after_stat(count) / sum(after_stat(count)) * 100,
      fill = Contains.super), alpha = 0.6) +
  facet_wrap(~info2, scales = "free") +
  labs(
    x = "value", 
    y = "frequency", 
    title = "Bar graph of superstition due to occupation,gender, and country"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 5.2.3. not-percentaged
```{r}
# Pivot the dataset and plot
top_filtered %>%
  pivot_longer(cols = c(Country, Occupation, Gender), names_to = "info2", values_to = "value2") %>%
  ggplot() +
  geom_bar(
    mapping = aes(
      x = value2, 
      fill = Contains.super), alpha = 0.6) +
  facet_wrap(~info2, scales = "free") +
  labs(
    x = "value", 
    y = "frequency", 
    title = "Bar graph of superstition due to occupation,gender, and country"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 5.3 Mosaic Graph
```{r}
# Filter out rows with missing values in the relevant columns
train_imputed <- top_filtered %>% 
  filter(!is.na(Gender) & !is.na(Occupation) & !is.na(Contains.super))

# Convert necessary columns to factors
train_imputed <- train_imputed %>%
  mutate(Gender = as.factor(Gender),
         Occupation = as.factor(Occupation),
         Contains.super = as.factor(Contains.super))
```
```{r}
# Create the mosaic plot
train_imputed %>% 
  ggplot() +
  geom_mosaic(aes(x = product(Gender, Occupation), fill = Gender)) +
  facet_grid(. ~ Contains.super, scales = "free") +
  labs(x = "Occupation", 
       y = "Gender", 
       title = "Mosaic Plot of Superstition") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

# 6. Logistic Regression

### 6.1. Women LR 

## 6.1.1. Women-Manipulate the data
```{r}
# Count and get the top 10 Occupations
top_occupations_2 <- test_age %>%
  count(Occupation) %>%
  top_n(10, wt = n) %>%
  pull(Occupation)

# Count and get the top 10 Countries
top_countries_2 <- test_age %>%
  count(Country) %>%
  top_n(10, wt = n) %>%
  pull(Country)

# Filter the train_age dataframe based on top Occupations and Countries
top_filtered_2 <- test_age %>%
  filter(Occupation %in% top_occupations_2) %>% 
  filter(Country %in% top_countries_2)
```
```{r}
# Filter out rows with missing values in the relevant columns
test_imputed <- top_filtered_2 %>% 
  filter(!is.na(Gender) & !is.na(Occupation) & !is.na(Contains.super))

# Convert necessary columns to factors
test_imputed <- test_imputed %>%
  mutate(Gender = as.factor(Gender),
         Occupation = as.factor(Occupation),
         Contains.super = as.factor(Contains.super))
```
```{r}
women_train_age <- train_imputed %>% 
    filter(Gender == 'Female') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))

women_test_age <- test_imputed %>% 
    filter(Gender == 'Female') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))
```

### 6.1.2. Modeling/pred - train
```{r}
w_train_model <- glm(Contains.super ~ Age.of.death, data = women_train_age, family = "binomial")
```
```{r}
w_train_preds <-  women_train_age %>% 
  add_predictions(w_train_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

w_train_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 6.1.3 Cross validation1 - train
```{r}
# Cost function for cross validation
cost <- function(obs, pred) {
  outcome <- pred > 0.5
  return(1 - (sum(obs == outcome) / length(obs)))
}

logistic_sv1 <- cv.glm(women_train_age, w_train_model, cost, K = 5)

logistic_sv1$delta
```

### 6.1.4. Modeling/pred - test
```{r}
w_test_model <- glm(Contains.super ~ Age.of.death + Occupation + Country, 
                    data = women_test_age, family = "binomial")
```
```{r}
w_test_preds <-  women_test_age %>% 
  add_predictions(w_test_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

w_test_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 6.1.5. Cross validation2 - test
```{r}
# Cost function for cross validation
cost <- function(obs, pred) {
  outcome <- pred > 0.5
  return(1 - (sum(obs == outcome) / length(obs)))
}

logistic_sv2 <- cv.glm(women_test_age, w_test_model, cost, K = 5)

logistic_sv2$delta
```

### 6.2. Men LR 

## 6.2.1.  Men-Manipulate the data
```{r}
men_train_age <- train_imputed %>% 
    filter(Gender == 'Male') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))

men_test_age <- test_imputed %>% 
    filter(Gender == 'Male') %>% 
    mutate(Shorter.lifespan = as.factor(Age.of.death < median(Age.of.death, na.rm = TRUE)))
```


### 6.2.2. Modeling/pred - train
```{r}
m_train_model <- glm(Contains.super ~ Age.of.death, data = men_train_age, family = "binomial")
```
```{r}
m_train_preds <-  men_train_age %>% 
  add_predictions(m_train_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

m_train_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 6.2.3. Cross validation3 - train
```{r}
logistic_sv3 <- cv.glm(men_train_age, m_train_model, cost, K = 5)

logistic_sv3$delta
```

### 6.2.4. Modeling/pred - test
```{r}
m_test_model <- glm(Contains.super ~ Age.of.death + Occupation + Country,
                    data = men_test_age, family = "binomial")
```
```{r}
m_test_preds <-  men_test_age %>% 
  add_predictions(m_test_model, type = "response") %>% 
  mutate(
    outcome = if_else(condition = pred > 0.5,
                      true = 1,
                      false = 0)
  )

m_test_preds %>% 
  mutate(
    correct = if_else(
      condition = Contains.four == outcome,
      true = 1,
      false = 0
    )
  ) %>% 
  summarize(
    total_correct = sum(correct),
    accuracy = total_correct/n()
  )
```

### 6.2.5. Cross validation3 - test
```{r}
logistic_sv4 <- cv.glm(men_test_age, m_test_model, cost, K = 5)

logistic_sv4$delta
```

# 7. Conclusion
> You can write conlcusion here.
